Title

### 2015-07-26 main.tex
minor tweaks

Computational biologists answer biological and biomedical 
questions by using computation in support of---or in place of---laboratory 
procedures, in hopes of obtaining more accurate answers at a greatly reduced 
cost.
The past two decades have seen unprecedented technological progress with regard
to generating biological data; next-generation sequencing, mass spectroscopy,
microarrays, nuclear magnetic resonance, and other high-throughput approaches
have led to an explosion of data.
This explosion of data is a mixed blessing.
On the one hand, the scale and scope of data should allow new insights into
genetic and infectious diseases, cancer, basic biology, and even human migration
patterns.
On the other hand, researchers are generating datasets so massive that it has 
become difficult to analyze them to discover patterns that give clues to the 
underlying biological processes.

Certainly, computers are getting faster and more economical; the amount of 
processing available per dollar of compute hardware is more or less doubling 
every year or two; a similar claim can be made about storage capacity 
(Figure~\ref{fig:growth}).
In 2002, when the first human genome was sequenced, the growth in computing 
power was still matching the growth rate of genomic data.
However, the sequencing technology used for the Human Genome Project--Sanger
sequencing--was supplanted around 2004, with the advent of what is now known as 
next-generation sequencing.
The material costs to sequence a genome have plummeted in the past
decade, to the point where a whole human genome can be sequenced for less than
one thousand U.S. Dollars.
As a result, the amount of genomic data available to 
researchers is increasing by a factor of 10 every year.

@@

Protein structure data, primarily determined through X-ray crystallography or
nuclear magnetic resonance (NMR), describes the coordinates of each atom, in 
Angstroms, of a protein in three-dimensional space.
This information is important to understanding how a protein functions, because
proteins interact physically with other molecules.
Protein structure data are commonly stored in the Protein Data Bank (PDB); 
each PDB entry has an associated resolution indicating the accuracy of the
structural description.
Thus a PDB entry at its most basic level contains a list of atoms, along with
which amino acid they belong to and their spatial coordinates as floating-point
numbers.

>>>>

Computational biologists answer biological and biomedical 
questions by using computation in support of---or in place of---laboratory 
procedures, in hopes of obtaining more accurate answers at a greatly reduced 
cost.
The past two decades have seen unprecedented technological progress with regard
to generating biological data; next-generation sequencing, mass spectroscopy,
microarrays, nuclear magnetic resonance spectroscopy, and other high-throughput 
approaches have led to an explosion of data.
However, this explosion of data is a mixed blessing.
On the one hand, the scale and scope of data should allow new insights into
genetic and infectious diseases, cancer, basic biology, and even human migration
patterns.
On the other hand, researchers are generating datasets so massive that it has 
become difficult to analyze them to discover patterns that give clues to the 
underlying biological processes.

Certainly, computers are getting faster and more economical; the amount of 
processing available per dollar of compute hardware is more or less doubling 
every year or two; a similar claim can be made about storage capacity 
(Figure~\ref{fig:growth}).
In 2002, when the first human genome was sequenced, the growth in computing 
power was still matching the growth rate of genomic data.
However, the sequencing technology used for the Human Genome Project--Sanger
sequencing--was supplanted around 2004, with the advent of what is now known as 
next-generation sequencing.
The material costs to sequence a genome have plummeted in the past
decade, to the point where a whole human genome can be sequenced for less than
one thousand U.S. Dollars.
As a result, the amount of genomic data available to 
researchers is increasing by a factor of ten every year.

@@

Protein structure data, primarily determined through X-ray crystallography or
nuclear magnetic resonance (NMR) spectroscopy, describes the coordinates of 
each atom, in Angstroms, of a protein in three-dimensional space.
This information is important to understanding how a protein functions, because
proteins interact physically with other molecules.
Protein structure data are commonly stored in the Protein Data Bank (PDB); 
each PDB entry has an associated resolution indicating the accuracy of the
structural description.
Thus a PDB entry at its most basic level contains a list of atoms, along with
which amino acid they belong to and their spatial coordinates as floating-point
numbers.
